---
title: "Covid91 vaccine study FinalFall2022"
author: "Sarah Medina"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ The vaccine has no effect on whether males will contract covid19

$H_1$ The vaccine effects the likely-hood a Male will contact covid19

## Graphical 

```{r}
barchartGC(~infected + treatment,data=M, type="percent")

```

The plot shows that of the males that contracted covid 19 over 60 percent of them received the placebo and about 30 percent received the vaccine. Of the males that did not contact covid19 50 percent received the placebo and 50 percent received the vaccine. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```
 The numerical results indicate that of the males who were infected with covid19, 68 percent received the placebo and 31 percent received the vaccine. The numerical results demonstrate the vaccine is effective for the male community since a larger percent of the males that got covid19 had the placebo and not the vaccine. The males who received the placebo were twice as likely to get covid19 than the males who got the vaccine. 

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
guysfisher <- fisher.test(table1)
guysfisher
```
The inferential results show that the p value is 2.364e-12. The p value is extremely low so we are able to reject the null hypothesis. Thus, the males who received the vaccine were less likely to contract covid19.


```{r}
effect <- guysfisher$p.value/(guysfisher$p.value +1)
effectivenessM = guysfisher$estimate/(guysfisher$estimate +1)*100
print(effectivenessM)                               
```



```{r}
low=(guysfisher$conf.int[1]/(guysfisher$conf.int[1] +1))*100
hi=(guysfisher$conf.int[2]/(guysfisher$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The effectiveness for males must be above 50 percent in order for the vaccine to be effective. As we can see the effective confidence Interval is from 63.4 to 73.35. Therefore, the vaccine is effective in males. Also the odds ratio for males is 69. The odds ratio indicates the effectiveness of the vaccine. Thus the vaccine is effective in males.


## Conclusion

We reject the null hypothesis that the vaccine has no effect on whether a male will contact covid19. This is because the p value provided by the fisher's exact test was extremely low. Therefore, the probability of there being no difference between the males who received the placebo and the males who received the vaccine is extremely unlikely. The numerical results indicated that of the males who received the vaccine 31 percent got covid19 and 68 percent of the males who received the placebo got covid19. Thus, the males who received the placebo were twice as likely to get covid19. If you are a male you should get the vaccine.  


## Females

$H_0$ The vaccine has no effect on whether Females will contract covid19

$H_1$ The vaccine effects the likely-hood a Female will contact covid19

## Graphical 

```{r}
barchartGC(~infected + treatment,data=F, type="percent")

```

The graphical results show that of the females who got covid19 about 60 percent had received the vaccine and about 40 percent received the placebo.


### Numerical Summaries  

```{r}
table1 <- xtabs(~infected + treatment, data=F)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that of the females that were infected with covid19 43 percent had taken the placebo and 57 percent had taken the vaccine. This suggest the vaccine may not be as effective for females as it was for males. It seems females are more likely to get covid19 if the take the vaccine. We will run a fisher exact test to confirm.

### Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesfish <- fisher.test(table1)
Femalesfish
```

The p value for females is extremely low. The p value is 0.005. Therefore, we are able to reject the null hypothesis. The vaccine does in fact effect the likelyhood of a feamle contracting covid19. Howerever, the graphical and numerical results show that females are more likely to contract covid19 if the take the vaccine.


```{r}
effect <- Femalesfish$p.value/(Femalesfish$p.value +1)
effectivenessF = Femalesfish$estimate/(Femalesfish$estimate +1)*100
print(effectivenessF)
```

```{r}
low=(Femalesfish$conf.int[1]/(Femalesfish$conf.int[1] +1))*100
hi=(Femalesfish$conf.int[2]/(Femalesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`


## LGBTQ



## Druggies


# Overall Results and Conclusions